---
title: "p8105_hw6_nh2706"
author: "Niklas Hess"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(corrplot)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

SOLUTIONS PROVIDED!

# Problem 2




# Problem 3

### Loading and cleaning the data for regression analysis

```{r}
birthweight_df = read_csv("./data/birthweight.csv") %>% 
  mutate(babysex = factor(if_else(babysex == 1, "male", "female")),
         frace = factor(recode(frace, '1' = "White", '2' = "Black", '3' = "Asian", 
                               '4' = "Puerto Rican", '8' = "Other", '9' = "Unknown")),
         mrace = factor(recode(mrace,'1' = "White", '2' = "Black", '3' = "Asian", 
                               '4' = "Puerto Rican", '8' = "Other", '9' = "Unknown")),
         malform = factor(recode(malform, '0' = "absent", '1' = "present")))

sum(is.na(birthweight_df))
```

### Proposing a regression model for birthweight

```{r}
birthweight_df %>%
  select(-babysex,-frace,-malform,-mrace,-pnumlbw,-pnumsga) %>%
  cor(method = "pearson") %>%
  corrplot(method = "number")
```
First, I create a correlation table. The two variables with the highest correlation with 'BWT' are 'BHEAD', 'GAWEEKS' and 'BLENGTH'. Therefore, I will use those three variables within my regression model. 

```{r}
bwt_reg1 = lm(bwt ~ blength + gaweeks + bhead, data = birthweight_df) 

bwt_reg1 %>%   
  broom::tidy() %>% 
  knitr::kable(digits = 2)
```
Second, I am constructing the linear model, including the aforementioned variables.

```{r}
birthweight_df %>% 
  modelr::add_residuals(bwt_reg1) %>% 
  modelr::add_predictions(bwt_reg1) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = .3) +
  labs(title = "Residuals against fitted values plot for Reg1 model",
       x = "Fitted Values",
       y = "Residuals") +
  theme(plot.title = element_text(hjust = 0.5))

```
Third, I am showing a plot of model residuals against fitted values. The plot shows that the linearity assumption likely holds.

### Comparing regression model to two others

#### Creating the two other models and checking their residuals vs. fitted values

Model #2: Using length at birth and gestational age as predictors (main effects only)
```{r}
bwt_reg2 = lm(bwt ~ blength + gaweeks, data = birthweight_df) 

bwt_reg2 %>%   
  broom::tidy() %>% 
  knitr::kable(digits = 2)

birthweight_df %>% 
  modelr::add_residuals(bwt_reg2) %>% 
  modelr::add_predictions(bwt_reg2) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = .3) +
  labs(title = "Residuals against fitted values plot for Reg2 model",
       x = "Fitted Values",
       y = "Residuals") +
  theme(plot.title = element_text(hjust = 0.5))
```


Model #3: Using head circumference, length, sex, and all interactions (including the three-way interaction)
```{r}
bwt_reg3 = lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data = birthweight_df) 

bwt_reg3 %>%   
  broom::tidy() %>% 
  knitr::kable(digits = 2)

birthweight_df %>% 
  modelr::add_residuals(bwt_reg3) %>% 
  modelr::add_predictions(bwt_reg3) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = .3) +
  labs(title = "Residuals against fitted values plot for Reg3 model",
       x = "Fitted Values",
       y = "Residuals") +
  theme(plot.title = element_text(hjust = 0.5))
```


#### Compariing using the cross-validated prediction error

```{r}
cv_df = 
  crossv_mc(birthweight_df, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)) %>% 
  mutate(
    model_bwt_reg1 = 
      map(train, ~lm(bwt ~ blength + gaweeks + bhead, data = birthweight_df)),
    model_bwt_reg2 = 
      map(train, ~lm(bwt ~ blength + gaweeks, data = birthweight_df)),
    model_bwt_reg3 = 
      map(train, ~lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data = birthweight_df))
  ) %>% 
  mutate(
    rmse_bwt_reg1 = map2_dbl(model_bwt_reg1, test, ~rmse(model = .x, data = .y)),
    rmse_bwt_reg2 = map2_dbl(model_bwt_reg2, test, ~rmse(model = .x, data = .y)),
    rmse_bwt_reg3 = map2_dbl(model_bwt_reg3, test, ~rmse(model = .x, data = .y))
  )
  
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_bwt_reg") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin() +
  labs(
    title = "RMSE values for three candidate model",
    x = "Model",
    y = "RMSEs"
  ) +
  theme(plot.title = element_text(hjust = 0.5))
```



