---
title: "p8105_hw6_nh2706"
author: "Niklas Hess"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(corrplot)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

SOLUTIONS PROVIDED!

# Problem 2

### Import and cleaning

Importing the data and cleaning it per instructions.
```{r}
raw_homicide_df = read_csv("./data/homicide-data.csv", na = c("","Unknown"))

homicide_df = raw_homicide_df %>%
  mutate(city_state = str_c(city, state, sep = ", ")) %>%
  filter(city_state != "Dallas, TX",
         city_state != "Phoenix, AZ",
         city_state != "Kansas City, MO",
         city_state != "Tulsa, AL",
         victim_race %in%  c("Black", "White")
        ) %>%
  mutate(solution = as.numeric(disposition == "Closed by arrest"),
         victim_age = as.numeric(victim_age),
         victim_race = as.numeric(victim_race == "White"), #White equals 1 and Black equals 0
         victim_sex = as.numeric(victim_sex == "Male")) #Female equals 0 and Male equals 1
```

### Baltimore Analysis

First, using the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors for Baltimore. Saving the output as an object.
```{r}
baltimore_df = homicide_df %>%
  filter(city == "Baltimore") %>%
  glm(solution ~ victim_age + victim_race + victim_sex, data = ., family = binomial()) %>%
  saveRDS(file = "data/baltimore_glm.rds")
```

Second, reading the object and showing the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims.
```{r}
readRDS(file = "data/baltimore_glm.rds") %>%
  broom::tidy(conf.int = TRUE) %>% 
  janitor::clean_names() %>%
  mutate(OR = exp(estimate),
         conf_low = exp(conf_low),
         conf_high = exp(conf_high)) %>%
  select(term, OR, conf_low, conf_high) %>% 
  filter(term == "victim_sex") %>%
  knitr::kable(digits = 2)
```

### All city analysis

Runing glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims.
```{r}
city_list = unique(homicide_df$city_state)

city_function = function(y){
  output = homicide_df %>%
    filter(city_state == y) %>%
    glm(solution ~ victim_age + victim_race + victim_sex, data = ., family = binomial()) %>%
    broom::tidy(conf.int = TRUE) %>% 
    janitor::clean_names() %>%
    mutate(OR = exp(estimate),
         conf_low = exp(conf_low),
         conf_high = exp(conf_high)) %>%
    filter(term == "victim_sex") %>%
    select(term, OR, conf_low, conf_high)
}

final_list = map(city_list, city_function)

df_final = data.frame(city_list[1],final_list[[1]])

names(df_final)[1] = "city_state"

for (i in 2:47) {
  df = data.frame(city_list[i],final_list[[i]])
  names(df)[1] = "city_state"
  df_final = rbind(df_final, df)
}

knitr::kable(df_final, digits = 2)
```

### Plotting results
Creating a plot that shows the estimated ORs and CIs for each city. Organizing cities according to estimated OR, and comment on the plot. Important to note that Male equals to 1, which means Female is the reference category.
```{r}
city_graph = df_final %>%
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf_low, ymax = conf_high)) +
  labs(
    title = "Estimated ORs and related CIs for each city",
    x = "City_State",
    y = "ORs") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

city_graph
```
Looking at the graph it becomes clear that the odd of solving a homicide comparing male victims to female victims varies widely. Two examples:

-  In NYC, the odds of solving a homicide for a male in NYC are 0.26 (CI: 0.13, 0.49) times the odds of solvings a homicide for females.
-   In Albuquerque, the odds of solving a homicide for a male in NYC are 1.77 (CI: 0.82, 3.76) times the odds of solvings a homicide for females. The 95% confidence interval includes 1, so likely not significant.

# Problem 3

### Loading and cleaning the data for regression analysis

```{r}
birthweight_df = read_csv("./data/birthweight.csv") %>% 
  mutate(babysex = factor(if_else(babysex == 1, "male", "female")),
         frace = factor(recode(frace, '1' = "White", '2' = "Black", '3' = "Asian", 
                               '4' = "Puerto Rican", '8' = "Other", '9' = "Unknown")),
         mrace = factor(recode(mrace,'1' = "White", '2' = "Black", '3' = "Asian", 
                               '4' = "Puerto Rican", '8' = "Other", '9' = "Unknown")),
         malform = factor(recode(malform, '0' = "absent", '1' = "present")))

sum(is.na(birthweight_df))
```

### Proposing a regression model for birthweight

```{r}
birthweight_df %>%
  select(-babysex,-frace,-malform,-mrace,-pnumlbw,-pnumsga) %>%
  cor(method = "pearson") %>%
  corrplot(method = "number")
```
First, I create a correlation table. The two variables with the highest correlation with 'BWT' are 'BHEAD', 'GAWEEKS' and 'BLENGTH'. Therefore, I will use those three variables within my regression model. 

```{r}
bwt_reg1 = lm(bwt ~ blength + gaweeks + bhead, data = birthweight_df) 

bwt_reg1 %>%   
  broom::tidy() %>% 
  knitr::kable(digits = 2)
```
Second, I am constructing the linear model, including the aforementioned variables.

```{r}
birthweight_df %>% 
  modelr::add_residuals(bwt_reg1) %>% 
  modelr::add_predictions(bwt_reg1) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = .3) +
  labs(title = "Residuals against fitted values plot for Reg1 model",
       x = "Fitted Values",
       y = "Residuals") +
  theme(plot.title = element_text(hjust = 0.5))

```
Third, I am showing a plot of model residuals against fitted values. The plot shows that the linearity assumption likely holds.

### Comparing regression model to two others

#### Creating the two other models and checking their residuals vs. fitted values

Model #2: Using length at birth and gestational age as predictors (main effects only)
```{r}
bwt_reg2 = lm(bwt ~ blength + gaweeks, data = birthweight_df) 

bwt_reg2 %>%   
  broom::tidy() %>% 
  knitr::kable(digits = 2)

birthweight_df %>% 
  modelr::add_residuals(bwt_reg2) %>% 
  modelr::add_predictions(bwt_reg2) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = .3) +
  labs(title = "Residuals against fitted values plot for Reg2 model",
       x = "Fitted Values",
       y = "Residuals") +
  theme(plot.title = element_text(hjust = 0.5))
```


Model #3: Using head circumference, length, sex, and all interactions (including the three-way interaction)
```{r}
bwt_reg3 = lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data = birthweight_df) 

bwt_reg3 %>%   
  broom::tidy() %>% 
  knitr::kable(digits = 2)

birthweight_df %>% 
  modelr::add_residuals(bwt_reg3) %>% 
  modelr::add_predictions(bwt_reg3) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = .3) +
  labs(title = "Residuals against fitted values plot for Reg3 model",
       x = "Fitted Values",
       y = "Residuals") +
  theme(plot.title = element_text(hjust = 0.5))
```


#### Compariing using the cross-validated prediction error

```{r}
cv_df = 
  crossv_mc(birthweight_df, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)) %>% 
  mutate(
    model_bwt_reg1 = 
      map(train, ~lm(bwt ~ blength + gaweeks + bhead, data = birthweight_df)),
    model_bwt_reg2 = 
      map(train, ~lm(bwt ~ blength + gaweeks, data = birthweight_df)),
    model_bwt_reg3 = 
      map(train, ~lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data = birthweight_df))
  ) %>% 
  mutate(
    rmse_bwt_reg1 = map2_dbl(model_bwt_reg1, test, ~rmse(model = .x, data = .y)),
    rmse_bwt_reg2 = map2_dbl(model_bwt_reg2, test, ~rmse(model = .x, data = .y)),
    rmse_bwt_reg3 = map2_dbl(model_bwt_reg3, test, ~rmse(model = .x, data = .y))
  )
  
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_bwt_reg") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin() +
  labs(
    title = "RMSE values for three candidate model",
    x = "Model",
    y = "RMSEs"
  ) +
  theme(plot.title = element_text(hjust = 0.5))
```

The violin graph above shows that model 1 and model 3 have a very similar RMSE score. As such, it would require some further investigation to decide which model is the best. Nevertheless, model 2 has a very high RMSE score and can be excluded.


# END